### 2023-08-18 sections/1-introduction.tex
struggles with intro and methods

Write intro \dots

@@

This paper focuses on:
\begin{itemize}
    \item Improvements and extensions of CHESS.
    \item Accelerated $\rho$-nearest and k-nearest neighbors search.
    \item Entropy-scaling (?? is it though?) compression using the hierarchical clustering.
    \item Proof for how cluster-radii, and thus compression deltas, fall as we descend along the tree.
    \item Information theoretic bounds on compression performance. We compress along the dimensionality of data rather than the cardinality of data.
    \item A variety of distance functions including euclidean, manhattan, cosine, jaccard, wasserstein, hamming, levenshtein, needleman-wunch, and any user-defined function.
\end{itemize}

Comparisons for:
\begin{itemize}
    \item Search time vs FALCONN and other in ann-benchmarks suite.
    \item Compression performance of CAKES+gzip vs just gzip.
\end{itemize}

>>>>

Often dubbed "the Big Data explosion", rapidly advancing technologies have enabled researchers to collect complex data on an unprecedented scale. 
For example, the Sloan Digital Sky Survey (SDSS) has collected data on over 500 million celestial objects~\cite{alam2015eleventh}. Or, as another 
example, the National Center for Biotechnology Information (NCBI) has collected over 2.25 million ribosomal DNA sequences~\cite{silva2014silva}. Finally, 
as a third example, the Large Hadron Collider (LHC) has collected over 100 petabytes of data~\cite{evans2008lhc}. Or, as a fourth example, the 
National Oceanic and Atmospheric Administration (NOAA) has collected over 20 petabytes of data~\cite{noaa2019noaa}. 
In many cases, the influx of data outpaces the rate of improvement in computing capacity predicted by Moore's Law~\cite{moore1965cramming}, indicating 
that computer hardware will not "catch up" to our computational needs in the near future. As a result, analyzing and learning from these large datasets 
requires better algorithms.

The ability to perform similarity search on these datasets enables a variety of applications, including recommendation and classification systems. 
Additionally, with the rise of large language models such as GPT-3~\cite{brown2020language}, similarity search can become a useful tool for 
understanding and verifying the validity of model outputs. 
As the sizes and dimensionalities of datasets have grown, efficient and accurate similarity search has become excessively challenging; 
state-of-the-art algorithms exhibit a steep recall versus speed tradeoff~\cite{ishaq2019clustered}.
In particular, k-nearest neighbors (kNN) search is one of the most pervasive supervised classification methods in use~\cite{fix1952discriminatory, cover1967nearest}.

Naive implementations of kNN, whose asymptotic complexity is a function of the dataset's cardinality, 
prove prohibitively slow for large datasets. While algorithms for kNN on large datasets exist, most of them are approximate and fail to 
exploit the inherent structure and redundancy often present in big datasets. While approximate search is sufficient for many applications, the 
need for efficient, exact search remains.

Recent approaches to tackle the exponential growth of data include locality-sensitive hashing~\cite{bingham2001random}, 
clever indexing techniques such as the FM Index~\cite{ferragina2005indexing}, and entropy-scaling search~\cite{yu2015entropy}. 
Entropy-scaling search is a paradigm for similarity search that exploits the inherent structure and redundancy often present in big datasets.
Importantly, as suggested by their name, entropy-scaling search algorithms have asymptotic complexity that scales with geometric properties of the dataset,
rather than its cardinality. In 2019, we introduced CHESS (CLustered Hierarchical Entropy-Scaling Search), which extended entropy-scaling $\rho$-nearest 
neighbors search to a hierarchical clustering approach. In this paper, we introduce CLAM (Clustered Learning of Approximate Manifolds), and CAKES (CLAM-Accelerated K-nearest-neighbor 
Entropy-scaling Search). CLAM, like CHESS, is a hierarchical clustering algorithm (figure out and explain improvements upon CHESS present in CLAM if there are any algorithmic ones)
that can facilitate a variety of algorithms for big data. Using the tree constructed by CLAM, CAKES builds upon the capabilities of CHESS to perform 
k-nearest neighbors search in addition to $\rho$-nearest neighbors search.

At some point, need to address the fact that exact search is important for small k and approxiate 
is maybe more ideal for large k. Try to find citation for this.

@@

This paper focuses on:
\begin{itemize}
    \item Improvements and extensions of CHESS.
    \item Accelerated $\rho$-nearest and k-nearest neighbors search.
    \item Proof for how cluster-radii fall as we descend along the tree.
    \item A variety of distance functions including euclidean, manhattan, cosine, jaccard, hamming, levenshtein, needleman-wunch, and any user-defined function.
\end{itemize}

Comparisons for:
\begin{itemize}
    \item Search time vs FAISS and FALCONN and other in ann-benchmarks suite.
\end{itemize}

### 2023-08-18 sections/3-methods.tex
struggles with intro and methods

\subsection{Partition}
\label{subsec:methods:partition}

Partition algorithm for building the tree.
Lift this from the CHAODA paper.

Exact partition with $\mathcal{O}(n^2)$ cost vs approximate partition using $\sqrt{n}$ seeds to achieve $\mathcal{O}(n)$ cost.
Building the exact tree costs $\mathcal{O}(n^2 \log n)$ vs approximate tree for $\mathcal{O}(n \log n)$.

@@

\subsubsection{Expanding Threshold}
\label{subsubsec:methods:knn-search:expanding-threshold}

Let $candidates$ be a priority queue of clusters sorted by non-increasing $\delta_{min}$. Initialize $candidates$ with the root cluster.
Let $hits$ be a second priority queue of points sorted by non-decreasing distance from the query. This queue starts empty.

While $hits$ has fewer than $k$ clusters or the $\delta$ of the top priority hit is greater 
than the $\delta_{min}$ of the top priority candidate, do the following:
\begin{enumerate}
\item While the top priority candidate isn't a leaf, replace the top priority 
candidate with its children.
\item Remove the top priority candidate from $candidates$ and add all of its 
points to $hits$. 
\item Remove points from $hits$ until $hits$ has $k$ points. 
\end{enumerate}

Extract $hits$.

\begin{algorithm} % enter the algorithm environment
\caption{Expanding Threshold} % give the algorithm a caption
\label{alg:expanding_threshold} % and a label for \ref{} commands later in the document
\begin{algorithmic}[1] % enter the algorithmic environment
    \REQUIRE $tree$, $query$, $k$
    \STATE $candidates \leftarrow$ priority queue
    \STATE $candidates.push(tree.root)$
    \STATE $hits \leftarrow$ priority queue
    \WHILE{$|hits| < k \lor hits.peek().\delta > candidates.peek().\delta_{min}$}
        \WHILE{$!candidates.peek().isLeaf$}
            \STATE $[l, r] \leftarrow candidates.pop().children()$
            \STATE $candidates.push([l, r])$
        \ENDWHILE
        \STATE $leaf \leftarrow candidates.pop()$
        \STATE $hits.push(leaf)$
        \WHILE{$|hits| > k$}
            \STATE $hits.pop()$
        \ENDWHILE
    \ENDWHILE
    \STATE Return $hits$
\end{algorithmic}
\end{algorithm}

\subsubsection{Sieve V1}
\label{subsubsec:methods:knn-search:sieve-v1}
Let $candidates$ be a vector of $clusters$s sorted by non-decreasing $\delta_{max}$. Initialize $candidates$ with  
the root cluster. 
Let $hits$ be a priority queue of size $k$ containing points sorted by non-decreasing distance from the query. This queue starts empty.

In this variant of KNN, we repeatedly calculate a threshold distance $\tau$ such that no cluster with a $\delta_{minmax}$ less than $\tau$ 
can contain one of the $k$ nearest neighbors.

Repeat the following: 
\begin{enumerate}
\item Find the cluster $C_{\tau}$ in $candidates$ with the smallest $\delta_{max}$ such that no cluster whose $\delta_{min}$ is greater than $C_{\tau}$'s $\delta_{max}$ can contain one of the $k$ nearest neighbors.
\item Let the threshold $\tau$ be the $\delta_{max}$ of $C_{\tau}$.
\item Remove from $hits$ any points which are a distance greater than $\tau$ from the query. 
\item If no cluster in $candidates$ has a $\delta_{max}$ less than or equal to $\tau$, break.
\item Remove from $candidates$ any cluster whose $\delta_{min}$ is greater than $\tau$.
\item Transfer from $candidates$ to $hits$ all the points in any cluster whose cardinality is less than $k$ or who is a leaf. 
\item Replace all non-leaf clusters in $candidates$ with their children. 
\end{enumerate}

At this stage, if any clusters remain in $candidates$, we add all their points to $hits$. 
Extract $hits$.

\begin{algorithm} % enter the algorithm environment
    \caption{Sieve V1} % give the algorithm a caption
    \label{alg:sieve_v1} % and a label for \ref{} commands later in the document
    \begin{algorithmic}[2] % enter the algorithmic environment
        \REQUIRE $tree$, $query$, $k$
        \STATE $candidates \leftarrow$ priority queue
        \STATE $candidates.push(tree.root)$
        \STATE $hits \leftarrow$ priority queue
        \WHILE{$|hits| < k \lor hits.peek().\delta > candidates.peek().\delta_{min}$}
            \WHILE{$!candidates.peek().isLeaf$}
                \STATE $[l, r] \leftarrow candidates.pop().children()$
                \STATE $candidates.push([l, r])$
            \ENDWHILE
            \STATE $leaf \leftarrow candidates.pop()$
            \STATE $hits.push(leaf)$
            \WHILE{$|hits| > k$}
                \STATE $hits.pop()$
            \ENDWHILE
        \ENDWHILE
        \STATE Return $hits$
    \end{algorithmic}
    \end{algorithm}

@@

\subsection{Compression}
\label{subsec:methods:compression}

Encode and Decode algorithms.
These could be, and probably will be, different for each $(X, f)$ pairing.

Asymptotic complexity analysis for:
\begin{enumerate}[i.]
    \item encode (novel contribution)
    \item serialize
    \item compress
    \item decompress
    \item deserialize
    \item decode (novel contribution)
\end{enumerate}

% discrete vs continuous, and small alphabet vs large alphabet in discrete

>>>>

\subsection{Dataset and Distance Function}
\label{subsec:methods:dataset-and-distance-function}

We start with a dataset $\textbf{X} = \{x_1 \dots x_n\}$ with $n$ points and a distance function $f : (\textbf{X}, \textbf{X}) \mapsto \mathbb{R}^+$.
Given two points from the dataset, the distance function deterministically returns a non-negative real number. We require that the distance function be symmetric ($f(x, y) = f(y, x) \forall x, y \in \textbf{X}$) and 
that the distance between two points $x$ and $y$ be zero if and only if $x = y$. When, in addition to these constraints, we also require that the distance function obey 
the triangle inequality (i.e., when the distance function is a metric), we can guarantee that our search algorithms have perfect recall.

CLAM assumes the ``manifold hypothesis''~\cite{fefferman2016testing}, which states that datasets collected from constrained generating phenomena that are embedded in a high-dimensional space typically only occupy a low-dimensional manifold in that space.
In other words, we assume that the dataset is embedded in a $D$-dimensional space, but that the data only occupy a $d$-dimensional manifold, where $d \ll D$. While we often Euclidean notions, such as voids and volumes, to talk about geometric and topological 
properties of clusters and of the manifold, CLAM does not rely on such notions; they serve merely as a convenient, more intuitive way to talk about the underlying mathematics.

\subsection{Clustering}
\label{subsec:methods:clustering}

We start by building a divisive hierarchical clustering of the dataset with CLAM, using a 
similar recursive procedure as outlined in CHESS~\cite{ishaq2019clustered}, but with better selection of 
poles. For a cluster with $m$ points, we begin by computing a 
random sample of $\sqrt m$ points. We then compute the geometric median of this sample, which we call the 
cluster's center. We then compute the distance from the center to each point in the sample of $\sqrt m$ points. 
The point $l$ which is furthest from the center is designated the left pole, and the point $r$ which is furthest
from $l$ is designated the right pole. We then partition the cluster into a left child and a right child, where the 
left child contains all points in the cluster which are closer to $l$ than to $r$, and the right child contains all 
points in the cluster which are closer to $r$ than to $l$. Without loss of generality, we assign to the left child 
those points which are equidistant from $l$ and $r$. Starting from a root-cluster containing the entire dataset, we 
repeat this procedure until each leaf contains only one datum or until some other user-specified stopping criterion 
is met.

\begin{algorithm} % enter the algorithm environment
\caption{Partition} % give the algorithm a caption
\label{alg:partition} % and a label for \ref{} commands later in the document
\begin{algorithmic}[1] % enter the algorithmic environment
    \REQUIRE $cluster$
    \STATE $m \leftarrow \lfloor \sqrt{|cluster.points|} \rfloor$
    \STATE $seeds \leftarrow m$ random points from $cluster.points$
    \STATE $c \leftarrow$ geometric median of $seeds$
    \STATE $l \leftarrow \argmax d(c,x) \ \forall \ x \in cluster.points$
    \STATE $r \leftarrow \argmax d(l,x) \ \forall \ x \in cluster.points$
    \STATE $left \leftarrow \{x | x \in cluster.points \land d(l,x) \le d(r,x)\}$
    \STATE $right \leftarrow \{x | x \in cluster.points \land d(r,x) < d(l,x)\}$
    \IF{$|left| > 1$}
        \STATE Partition($left$)
    \ENDIF
    \IF{$|right| > 1$}
        \STATE Partition($right$)
    \ENDIF
\end{algorithmic}
\end{algorithm}

\subsubsection {Proof of Eventually Decreasing Cluster Radii}
\subsection{Dataset and Distance Function}
\label{subsec:methods:dataset-and-distance-function}

We start with a dataset $\textbf{X} = \{x_1 \dots x_n\}$ with $n$ points and a distance function $f : (\textbf{X}, \textbf{X}) \mapsto \mathbb{R}^+$.
Given two points from the dataset, the distance function deterministically returns a non-negative real number. We require that the distance function be symmetric ($f(x, y) = f(y, x) \forall x, y \in \textbf{X}$) and 
that the distance between two points $x$ and $y$ be zero if and only if $x = y$. When, in addition to these constraints, we also require that the distance function obey 
the triangle inequality (i.e., when the distance function is a metric), we can guarantee that our search algorithms have perfect recall.

CLAM assumes the ``manifold hypothesis''~\cite{fefferman2016testing}, which states that datasets collected from constrained generating phenomena that are embedded in a high-dimensional space typically only occupy a low-dimensional manifold in that space.
In other words, we assume that the dataset is embedded in a $D$-dimensional space, but that the data only occupy a $d$-dimensional manifold, where $d \ll D$. While we often Euclidean notions, such as voids and volumes, to talk about geometric and topological 
properties of clusters and of the manifold, CLAM does not rely on such notions; they serve merely as a convenient, more intuitive way to talk about the underlying mathematics.

\subsection{Clustering}
\label{subsec:methods:clustering}

We start by building a divisive hierarchical clustering of the dataset with CLAM, using a 
similar recursive procedure as outlined in CHESS~\cite{ishaq2019clustered}, but with better selection of 
poles. For a cluster with $m$ points, we begin by computing a 
random sample of $\sqrt m$ points. We then compute the geometric median of this sample, which we call the 
cluster's center. We then compute the distance from the center to each point in the sample of $\sqrt m$ points. 
The point $l$ which is furthest from the center is designated the left pole, and the point $r$ which is furthest
from $l$ is designated the right pole. We then partition the cluster into a left child and a right child, where the 
left child contains all points in the cluster which are closer to $l$ than to $r$, and the right child contains all 
points in the cluster which are closer to $r$ than to $l$. Without loss of generality, we assign to the left child 
those points which are equidistant from $l$ and $r$. Starting from a root-cluster containing the entire dataset, we 
repeat this procedure until each leaf contains only one datum or until some other user-specified stopping criterion 
is met.

\begin{algorithm} % enter the algorithm environment
\caption{Partition} % give the algorithm a caption
\label{alg:partition} % and a label for \ref{} commands later in the document
\begin{algorithmic}[1] % enter the algorithmic environment
    \REQUIRE $cluster$
    \STATE $m \leftarrow \lfloor \sqrt{|cluster.points|} \rfloor$
    \STATE $seeds \leftarrow m$ random points from $cluster.points$
    \STATE $c \leftarrow$ geometric median of $seeds$
    \STATE $l \leftarrow \argmax d(c,x) \ \forall \ x \in cluster.points$
    \STATE $r \leftarrow \argmax d(l,x) \ \forall \ x \in cluster.points$
    \STATE $left \leftarrow \{x | x \in cluster.points \land d(l,x) \le d(r,x)\}$
    \STATE $right \leftarrow \{x | x \in cluster.points \land d(r,x) < d(l,x)\}$
    \IF{$|left| > 1$}
        \STATE Partition($left$)
    \ENDIF
    \IF{$|right| > 1$}
        \STATE Partition($right$)
    \ENDIF
\end{algorithmic}
\end{algorithm}

\subsubsection {Proof of Eventually Decreasing Cluster Radii}
\label{subsubsec:methods:clustering:proof-of-eventually-decreasing-cluster-radii}

\subsubsection {Complexity}
\label{subsubsec:methods:clustering:complexity}

Clustering: Exact partition with $\mathcal{O}(n^2)$ cost vs approximate partition using $\sqrt{n}$ seeds to achieve $\mathcal{O}(n)$ cost.
Building the exact tree costs $\mathcal{O}(n^2 \log n)$ vs approximate tree for $\mathcal{O}(n \log n)$.

@@

\subsubsection{Expanding Threshold}
\label{subsubsec:methods:knn-search:expanding-threshold}
In this variant of KNN, we repeatedly

Let $Q$ be a priority queue of clusters sorted by non-increasing $\delta_{min}$. Initialize $Q$ with the root cluster.
Let $P$ be a second priority queue of points sorted by non-decreasing distance from the query. This queue starts empty.

While $P$ has fewer than $k$ clusters or $\delta$ of the top priority element in $P$ is greater 
than $\delta_{min}$ of the top priority element in $Q$, do the following:
\begin{enumerate}
\item While the top priority element $q$ in $Q$ isn't a leaf, replace $q$ with its children.
\item Remove $q$ from $Q$ and add all of its points to $P$. 
\item Remove points from $P$ until $P$ has $k$ points. 
\end{enumerate}

Return $P$.

\begin{algorithm} % enter the algorithm environment
\caption{Expanding Threshold} % give the algorithm a caption
\label{alg:expanding_threshold} % and a label for \ref{} commands later in the document
\begin{algorithmic}[1] % enter the algorithmic environment
    \REQUIRE $tree$, $query$, $k$
    \STATE $Q \leftarrow$ priority queue
    \STATE $Q.push(tree.root)$
    \STATE $P \leftarrow$ priority queue
    \WHILE{$|P| < k \lor P.max.\delta > Q.min.\delta_{min}$}
        \WHILE{$!Q.min.isLeaf$}
            \STATE $[l, r] \leftarrow Q.extractMin().children()$
            \STATE $Q.push([l, r])$
        \ENDWHILE
        \STATE $leaf \leftarrow Q.extractMin()$
        \STATE $P.push(leaf)$
        \WHILE{$|P| > k$}
            \STATE $P.extractMax()$
        \ENDWHILE
    \ENDWHILE
    \STATE Return $P$
\end{algorithmic}
\end{algorithm}

\subsubsection{Sieve V1}
\label{subsubsec:methods:knn-search:sieve-v1}

In this variant of KNN, we repeatedly calculate a threshold distance $\tau$ such that no cluster with a $\delta_{min}$ greater than $\tau$ 
can contain one of the $k$ nearest neighbors. Search terminates when we have a threshold containing exactly $k$ points.

Let $Q$ be a vector of clusters sorted by non-decreasing $\delta_{max}$. Initialize $Q$ with  
the root cluster. 
Let $P$ be a priority queue of size $k$ containing points sorted by non-decreasing distance from the query. This queue starts empty.

Repeat the following: 
\begin{enumerate}
\item Find the cluster $C_{\tau}$ in $Q$ with the smallest $\delta_{max}$ such that no cluster whose $\delta_{min}$ is greater than $C_{\tau}$'s $\delta_{max}$ can contain one of the $k$ nearest neighbors.
\item Let the threshold $\tau$ be the $\delta_{max}$ of $C_{\tau}$.
\item Remove from $P$ any points which are a distance greater than $\tau$ from the query. 
\item If no cluster in $Q$ has a $\delta_{max}$ less than or equal to $\tau$, break.
\item Remove from $Q$ any cluster whose $\delta_{min}$ is greater than $\tau$.
\item Transfer from $Q$ to $P$ all the points in any cluster whose cardinality is less than $k$ or who is a leaf. 
\item Replace all non-leaf clusters in $Q$ with their children. 
\end{enumerate}

At this stage, if any clusters remain in $Q$, we add all their points to $P$. 
Extract $P$.

\begin{algorithm} % enter the algorithm environment
    \caption{Sieve V1} % give the algorithm a caption
    \label{alg:sieve_v1} % and a label for \ref{} commands later in the document
    \begin{algorithmic}[2] % enter the algorithmic environment
        \REQUIRE $tree$, $query$, $k$
        \STATE $Q \leftarrow$ priority queue
        \STATE $Q.push(tree.root)$
        \STATE $P \leftarrow$ priority queue
        \WHILE{$|P| < k \lor P.max.\delta > Q.max.\delta_{min}$}
            \WHILE{$!Q.extractMax().isLeaf$}
                \STATE $[l, r] \leftarrow Q.extractMax().children()$
                \STATE $Q.push([l, r])$
            \ENDWHILE
            \STATE $leaf \leftarrow Q.extractMax()$
            \STATE $P.push(leaf)$
            \WHILE{$|P| > k$}
                \STATE $P.extractMax()$
            \ENDWHILE
        \ENDWHILE
        \STATE Return $P$
    \end{algorithmic}
    \end{algorithm}

@@

\label{subsubsec:methods:clustering:proof-of-eventually-decreasing-cluster-radii}

\subsubsection {Complexity}

\subsubsection{Expanding Threshold}
\label{subsubsec:methods:knn-search:expanding-threshold}
In this variant of KNN, we repeatedly

Let $Q$ be a priority queue of clusters sorted by non-increasing $\delta_{min}$. Initialize $Q$ with the root cluster.
Let $P$ be a second priority queue of points sorted by non-decreasing distance from the query. This queue starts empty.

While $P$ has fewer than $k$ clusters or $\delta$ of the top priority element in $P$ is greater 
than $\delta_{min}$ of the top priority element in $Q$, do the following:
\begin{enumerate}
\item While the top priority element $q$ in $Q$ isn't a leaf, replace $q$ with its children.
\item Remove $q$ from $Q$ and add all of its points to $P$. 
\item Remove points from $P$ until $P$ has $k$ points. 
\end{enumerate}

Return $P$.

\begin{algorithm} % enter the algorithm environment
\caption{Expanding Threshold} % give the algorithm a caption
\label{alg:expanding_threshold} % and a label for \ref{} commands later in the document
\begin{algorithmic}[1] % enter the algorithmic environment
    \REQUIRE $tree$, $query$, $k$
    \STATE $Q \leftarrow$ priority queue
    \STATE $Q.push(tree.root)$
    \STATE $P \leftarrow$ priority queue
    \WHILE{$|P| < k \lor P.max.\delta > Q.min.\delta_{min}$}
        \WHILE{$!Q.min.isLeaf$}
            \STATE $[l, r] \leftarrow Q.extractMin().children()$
            \STATE $Q.push([l, r])$
        \ENDWHILE
        \STATE $leaf \leftarrow Q.extractMin()$
        \STATE $P.push(leaf)$
        \WHILE{$|P| > k$}
            \STATE $P.extractMax()$
        \ENDWHILE
    \ENDWHILE
    \STATE Return $P$
\end{algorithmic}
\end{algorithm}

\subsubsection{Sieve V1}
\label{subsubsec:methods:knn-search:sieve-v1}

In this variant of KNN, we repeatedly calculate a threshold distance $\tau$ such that no cluster with a $\delta_{min}$ greater than $\tau$ 
can contain one of the $k$ nearest neighbors. Search terminates when we have a threshold containing exactly $k$ points.

Let $Q$ be a vector of clusters sorted by non-decreasing $\delta_{max}$. Initialize $Q$ with  
the root cluster. 
Let $P$ be a priority queue of size $k$ containing points sorted by non-decreasing distance from the query. This queue starts empty.

Repeat the following: 
\begin{enumerate}
\item Find the cluster $C_{\tau}$ in $Q$ with the smallest $\delta_{max}$ such that no cluster whose $\delta_{min}$ is greater than $C_{\tau}$'s $\delta_{max}$ can contain one of the $k$ nearest neighbors.
\item Let the threshold $\tau$ be the $\delta_{max}$ of $C_{\tau}$.
\item Remove from $P$ any points which are a distance greater than $\tau$ from the query. 
\item If no cluster in $Q$ has a $\delta_{max}$ less than or equal to $\tau$, break.
\item Remove from $Q$ any cluster whose $\delta_{min}$ is greater than $\tau$.
\item Transfer from $Q$ to $P$ all the points in any cluster whose cardinality is less than $k$ or who is a leaf. 
\item Replace all non-leaf clusters in $Q$ with their children. 
\end{enumerate}

At this stage, if any clusters remain in $Q$, we add all their points to $P$. 
Extract $P$.

\begin{algorithm} % enter the algorithm environment
    \caption{Sieve V1} % give the algorithm a caption
    \label{alg:sieve_v1} % and a label for \ref{} commands later in the document
    \begin{algorithmic}[2] % enter the algorithmic environment
        \REQUIRE $tree$, $query$, $k$
        \STATE $Q \leftarrow$ priority queue
        \STATE $Q.push(tree.root)$
        \STATE $P \leftarrow$ priority queue
        \WHILE{$|P| < k \lor P.max.\delta > Q.max.\delta_{min}$}
            \WHILE{$!Q.extractMax().isLeaf$}
                \STATE $[l, r] \leftarrow Q.extractMax().children()$
                \STATE $Q.push([l, r])$
            \ENDWHILE
            \STATE $leaf \leftarrow Q.extractMax()$
            \STATE $P.push(leaf)$
            \WHILE{$|P| > k$}
                \STATE $P.extractMax()$
            \ENDWHILE
        \ENDWHILE
        \STATE Return $P$
    \end{algorithmic}
    \end{algorithm}

Clustering: Exact partition with $\mathcal{O}(n^2)$ cost vs approximate partition using $\sqrt{n}$ seeds to achieve $\mathcal{O}(n)$ cost.
Building the exact tree costs $\mathcal{O}(n^2 \log n)$ vs approximate tree for $\mathcal{O}(n \log n)$.
